---
title: "Homework 1"
author: Ashwini Varghese
output: github_document
date: "2022-09-22"
---

# Problem 1:

After installing the package with the penguin dataset, we will load it:

```{r beginning}
library(palmerpenguins)

data("penguins", package = "palmerpenguins")
```
After selecting the _penguins_ dataset from the environment pane, we can visually inspect it in the console. This dataset has contains information about different species of penguins. The variables included provide information about each penguin's _species, island, bill length, bill depth, flipper length, body mass, sex, and the year_.

This dataset has 344 rows, which we found by using the code `nrow(penguins)` and 8 columns, found using code `ncol(penguins)`.

To find the mean flipper length, we will first load the tidyverse package `library(tidyverse)` to be able to use the `pull` function to find the mean.

Using the code `mean(pull(penguins, var = flipper_length_mm), na.rm = TRUE)` we get the mean flipper length of **200.9152**.  


We will now load the `ggplot2` package and make a scatter plot with `bill_length_mm` as the _x_ variable and `flipper_length_mm` as the _y_ variable. We will also add color to distinguish the different `species`. 

```{r second_block}
library(ggplot2)

ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
```

Now we will export this plot to the directory.

```{r third_block}
ggsave("First_plot.pdf", height = 4, width = 4)
```


# Problem 2:

For problem 2, we will create a data frame of 10 observations and 4 variables.

```{r fourth_block}
library(tibble)

four_df = tibble (
  random = rnorm(10),
  vec_logic = random > 0,
  vec_chara = c("car", "boat", "truck", "plane", "jet", "cycle", "moped", "earth", "sea", "land"),
  vec_factor = factor(c("top", "middle", "bottom", "top", "middle", "bottom", "top", "middle", "bottom", "top"))
)

four_df
```

We will try to take the mean of each variable.

```{r averages, collapse = TRUE}
library(tidyverse)

mean(pull(four_df, random))
mean(pull(four_df, vec_logic))
mean(pull(four_df, vec_chara))
mean(pull(four_df, vec_factor))
```
We are able to get the mean of the variables `random` and `vec_logic`. However the other two variables, `vec_chara` and `vec_factor`, did not produce an average because the values are not logical nor numerical.

We can attempt to convert the logic, character, and factor variables to a numeric variables using the `as.numeric` functioN.

```{r conversion, eval = FALSE}
as.numeric(factor(c("car", "boat", "truck", "plane", "jet", "cycle", "moped", "earth", "sea", "land")))
as.numeric(factor(c("top", "middle", "bottom", "top", "middle", "bottom", "top", "middle", "bottom", "top")))
as.numeric(factor(c(random > 0)))
```
When we use this function, it converts the values in the character and factor variables into numeric values. This shows why we were unable to get the mean previously.

